---
title: "homework5"
author: "Sam Shifflett"
date: "4/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

3.a)
```{r}
library("tidyverse")
url <- "https://xdaiisu.github.io/ds202materials/data/ames.csv"
data <- readr::read_csv(url)
data <- separate(data, col="Store Location",into = c("Street Address","Lat","Lon"), sep="([\\(\\,\\)])")
data <- separate(data, col="Invoice/Item Number", into = c("INV","Item Number"), sep = "-")
head(data)
```



3.b)
```{r}
sapply(data,class)
#lat,lon,vendornumber,city(factor),vendorname(factor)
data[,c("Lat","Lon","Vendor Number", "Item Number")] = sapply(data[,c("Lat","Lon","Vendor Number", "Item Number")], as.numeric)
data[,c("City","Vendor Name")] = sapply(data[,c("City","Vendor Name")], as.factor)
```

3.c)
```{r}
data = separate(data=data,col=Date,into=c("Month","Day","Year"),sep="\\/")
data[,c("Month","Day","Year")] = sapply(data[,c("Month","Day","Year")],as.numeric)
head(data)
```
4.a)
```{r}
library(ggplot2)
ggplot(data, aes(x=Lat,y=Lon)) + geom_point() + ylim(-93.5,-93.8)
```

4.b)
```{r}
library(gridExtra)
names(data) <- make.names(names(data))
data <- transform(data, Category.Name=as.factor(Category.Name), group=as.factor(findInterval(Category.Name, c(1,20,40,60,80,100,124))))
data <- transform(data, Category.Name=as.factor(Category.Name), group=as.factor(findInterval(Category.Name, c(1,20,40,60,80,100,124))))
g1 = ggplot(subset(data,data$group==1), aes(Category.Name,y=Volume.Sold..Liters.)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8))
g2 = ggplot(subset(data,data$group==2), aes(Category.Name,y=Volume.Sold..Liters.)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8))
g3 = ggplot(subset(data,data$group==3), aes(Category.Name,y=Volume.Sold..Liters.)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8))
g4 = ggplot(subset(data,data$group==4), aes(Category.Name,y=Volume.Sold..Liters.)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8))
g5 = ggplot(subset(data,data$group==5), aes(Category.Name,y=Volume.Sold..Liters.)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8))
g6 = ggplot(subset(data,data$group==6), aes(Category.Name,y=Volume.Sold..Liters.)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8))
#grid.arrange(g1,g2,g3,g4,g5,g6,g7,ncol=2,nrow=4)
g1
g2
g3
g4
g5
g6
```


```{r}
filterdata = filter(data, Year == 2018 & City == "Ames")
names(filterdata) <- make.names(names(filterdata))
sumdaydata = filterdata %>% group_by(Month,Day) %>% summarize(Sales=sum(Sale..Dollars.))
ggplot(sumdaydata, aes(x=Day,y=Sales, color=Month)) + geom_point() + facet_wrap(sumdaydata$Month)
#Game Dates
#SEP 1 (SAT)
#SEP 8 (SAT)
#SEP 15 (SAT)
#SEP 22 (SAT) 
#SEP 29 (SAT)
#OCT 6 (SAT)
#OCT 13 (SAT)
#OCT 27 (SAT)
#NOV 3 (SAT)
#NOV 10 (SAT)
#NOV 17 (SAT)
#NOV 24 (SAT)
#DEC 1 (SAT) 
#DEC 28 (FRI)
```



```{r}
sumdaydataVolume = filterdata %>% group_by(Month,Day) %>% summarize(Volume.Sold=sum(Volume.Sold..Liters.))
ggplot(sumdaydataVolume, aes(x=Day,y=Volume.Sold, color=Month)) + geom_line() + facet_wrap(sumdaydata$Month)
```


By observing the two plots, it is easy to see that every Friday/Saturday liqour sales increase dramatically. Even during the dates of the football games, there seems to be a standard increase in sales and volume sold. I thought there would be more of a drastic increase during games, but that doesn't seem to be the case. I guess that the lure of a football game doesn't bring that drastric of an increase in sales to the liqour market here in ames. I have a feeling that this is because the people who travel here to see the games buy their alcohol elsewhere before arriving.